### 2010-12-16 docs/CrystalPlan_paper/paper.tex
Refs #182. All Pete's comments are in.

%------------------------------------------------------------------------------
% Template file for the submission of papers to IUCr journals in LaTeX2e
% using the iucr document class
% Copyright 1999-2009 International Union of Crystallography
% Version 1.4 (11 May 2009)
%------------------------------------------------------------------------------

\documentclass[final]{iucr}              % DO NOT DELETE THIS LINE

     %-------------------------------------------------------------------------
     % Information about the type of paper
     %-------------------------------------------------------------------------
     \paperprodcode{a000000}      % Replace with production code if known
     \paperref{xx9999}            % Replace xx9999 with reference code if known
     \papertype{FA}               % Indicate type of article
                                  %   FA - research papers (full article)
                                  %   SC - short communications
                                  %   LA - lead article
                                  %   FE - feature articles
                                  %   ST - structural communications
                                  %   XC - crystallization communications

     \paperlang{english}          % Can be english, french, german or russian
     %-------------------------------------------------------------------------
     % Information about journal to which submitted
     %-------------------------------------------------------------------------
     \journalcode{J}              % Indicate the journal to which submitted
                                  %   A - Acta Crystallographica Section A
                                  %   B - Acta Crystallographica Section B
                                  %   C - Acta Crystallographica Section C
                                  %   D - Acta Crystallographica Section D
                                  %   E - Acta Crystallographica Section E
                                  %   F - Acta Crystallographica Section F
                                  %   J - Journal of Applied Crystallography
                                  %   S - Journal of Synchrotron Radiation
          %--------------------------------------------------------------------
          % The following entries will be changed as required by editorial staff
          %--------------------------------------------------------------------
     \journalyr{2010}
     \journaliss{1}
     \journalvol{65}
     \journalfirstpage{000}
     \journallastpage{000}
     \journalreceived{0 XXXXXXX 0000}
     \journalaccepted{0 XXXXXXX 0000}
     \journalonline{0 XXXXXXX 0000}

\newif\ifdograph
% \dographtrue % DO keep the graphs.
\dographfalse % DONT put graphs - makes things faster

\usepackage{ifpdf}

\usepackage{graphicx}
\DeclareGraphicsExtensions{.eps,.ps,.bmp,.tif,.tiff,.tga}  
\graphicspath{{eps/}}

\usepackage{gensymb} %General symbols
\usepackage[version=3]{mhchem} %For chemical formulae
\usepackage{subfigure}

%quick angstrom define
\newcommand{\ang}{$\mathrm{\AA} $}
%\newcommand{\th}[0]{\superscript{th}}

\begin{document}                  % DO NOT DELETE THIS LINE

 %-------------------------------------------------------------------------
 % The introductory (header) part of the paper
 %-------------------------------------------------------------------------

\title{CrystalPlan: an Experiment Planning Tool for Crystallography}

\shorttitle{CrystalPlan: an Experiment Planning Tool}

 % Authors' names and addresses. Use \cauthor for the main (contact) author.
 % Use \author for all other authors. Use \aff for authors' affiliations.
 % Use lower-case letters in square brackets to link authors to their
 % affiliations; if there is only one affiliation address, remove the [a].

\cauthor[a]{Janik}{Zikovsky}{zikovskyjl@ornl.gov}{address if different from
\aff}
\author[a]{Peter}{Peterson}
\author[a]{Xiaoping}{Wang}
\author[a]{Matthew}{Frost}
\author[a]{Christina}{Hoffmann}

\aff[a]{Spallation Neutron Source, Oak Ridge National Laboratory, P.O. Box 2008
MS-6477, Oak Ridge, TN 37831-6477
\country{USA}}

 % Use \shortauthor to indicate an abbreviated author list for use in
 % running heads (you will need to uncomment it).

\shortauthor{Zikovsky et al.}

\maketitle                        % DO NOT DELETE THIS LINE

\begin{synopsis}
Describes CrystalPlan, a software program for the automatic creation of 
crystallography experiment plans for time-of-flight diffractometers.
\end{synopsis}

\begin{abstract}
Beam time at large x-ray and neutron scattering facilities is always at a premium.
The CrystalPlan program can calculate the data coverage of a crystal in reciprocal
space in a single-crystal diffraction time-of-flight experiment. CrystalPlan can 
help a user build an experiment plan that will acquire the most data possible, 
with sufficient coverage but limited redundancy, therefore increasing scientific 
productivity. 
A user friendly GUI including a 3D viewer, an automated coverage optimizer,
and an option to reorient the crystal for the measurement 
of selected hkls on specific detector positions 
are among its useful features. 
A sample use case of the program with the TOPAZ beamline at SNS will be
presented. 
\end{abstract}

%-------------------------------------------------------------------------
%-------------------------------------------------------------------------
%-------------------------------------------------------------------------
\section{Introduction}

The TOPAZ beamline at the Spallation Neutron Source (SNS) at Oak Ridge National
Lab (ORNL), currently finishing commissioning, is a neutron time-of-flight
Laue instrument with high wavelength resolution. It is designed to acquire
single-crystal diffraction data of small-to-moderate sized unit cells with  high
throughput. The final design calls for an array of 48 detectors covering a large
fraction of $2\pi$ steradians. However as of this writing, only 14 of the 48
detectors are installed, giving only a fraction of $2\pi$ steradian coverage. 
In order to have sufficient data to perform a fit, an experimenter will typically want
to measure $ > 85\%$ of the peaks within a certain range of d-spacings. 
However, wide wavelength band combined with TOPAZ's complex
geometry makes it very difficult to come up with an effective set of sample
orientations that will cover all the required peaks without unnecessary redundancy.

Existing diffraction experiment planning software programs (for example,
CrysAlis Pro from Oxford Diffraction; COSMO from Bruker AXS) were designed for
X-ray experiments where a single  incident wavelength and a large number of
sample orientations or sweeps are used to collect data. 
These programs are not suitable for TOPAZ, since as a time-of-flight instrument, 
it is able to collect data from a wide bandwidth of
incident neutrons, using a stationary crystal. 
Additionally, this means that relatively few sample orientations are needed in
order to complete a measurement: whereas an X-ray measurement might have
hundreds or thousands of individual frames, a TOPAZ run may have as few as 10.

The goal of CrystalPlan was to make an easy-to-use software tool to aid users in
planning their experiments. By quickly making an experimental plan that will be
sufficient for the needs of the scientist (enough peaks measured), but without
spending time needlessly measuring the same region in reciprocal space,
CrystalPlan will greatly increase the scientific productivity and throughput of
TOPAZ and other beamlines. Users will be able to measure more
samples or more fully explore parameters such as temperature and pressure.

%---------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------
\section{Architecture}

CrystalPlan is designed with these capabilities:

\begin{itemize}
  \item Load an instrument geometry.
  \item Calculate the orientation of a crystal at different goniometer settings.
  \item Calculate the coverage of these orientations in reciprocal space,
  in volume and for single peaks.
  \item Take crystal Bravais lattice symmetries into account.
  \item Optimize the experiment plan automatically.
  \item Compare predicted and actual measurement of peaks.
\end{itemize}

\subsection{Instrument Configuration}

CrystalPlan supports area detectors with flat faces and arbitrary rectangular
dimensions; this handles the detector types in use at SNS (Anger cameras and
8-pack tube detectors), but the software could easily be extended for more
complicated detector shapes as needed. The detector configuration is loaded from
a text file containing the location, size, orientation and number of pixels of
each detector. Other settings to be set for an instrument include the
detectable range of neutron/x-ray wavelength.

\subsection{Sample Orientation}

CrystalPlan is designed to be able to use a variety of sample orientation
goniometers. A base Goniometer class is subclassed for specific types of
goniometers, allowing a unified interface. This base class acts as a
\emph{universal goniometer},
which uses the conventional goniometer angles of
$\phi$, $\chi$, $\omega$ as defined in \cite{busing67}.
Each class inheriting from Goniometer can convert
the desired $\phi$, $\chi$, $\omega$ angles to the internal motor positions it requires. The API allows
for limited goniometers: for example, TOPAZ is currently equipped with a
goniometer where chi is fixed at 45 degrees. CrystalPlan handles this by
limiting the available sample orientation angles accessible to the user.

From an initial data set, the data analysis program ISAW \cite{Tao2006} is used
to calculate the the sample mounting orientation UB matrix, which is saved to a text file along with the lattice
parameters. CrystalPlan can load this UB matrix file and will use that UB matrix
in coverage calculations.

\subsection{Reciprocal Space Representation.}

The reciprocal space to be measured is divided into a cubic 3D matrix, the size
of which is determined by the user, who specifies the minimum d-spacing
($d_{min}$, in \ang) that he or she is interested in. The resolution of the
3D matrix is also specified (in \ang$^{-1}$), which determines the number of
points (voxels or volumetric pixels) to be modeled. Available machine memory
limits the resolution that can be achieved in practice.

Each point of the reciprocal space 3D matrix holds an integer representing the 
number of times that voxel of {\bf Q}-space (with {\bf Q} defined as $\frac{2\pi}{d}$) has
been measured -- this is the total coverage matrix. To speed up the
calculation of the total, a coverage matrix for each sample orientation is 
calculated and saved in memory. This means that the total coverage matrix can
be computed by simply adding multiple 3D matrices corresponding to each sample orientation.

\subsection{Calculating Coverage of One Sample Orientation}

For each sample orientation, a rotation matrix combining the sample's mounting
orientation (the U matrix loaded from ISAW) and the goniometer rotation matrices
($\Phi$, $\chi$, $\Omega$) is computed. The coverage in reciprocal space is
computed in this way: for each pixel in the face of a detector, the direction of
the scattered beam is known. Then, two incident beam vectors are considered,
corresponding to the $\lambda_{min}$ and $\lambda_{max}$ wavelength limits
of the detectors and/or source. The scattered beam direction and the two possible incident beam vectors
are used to to find two $\overrightarrow q$ vectors; these point in the
same direction. At the high wavelength we find $\overrightarrow q_{min}$, and
the low wavelength limit scattering off of $\overrightarrow q_{max}$. Since the
instrument has a wide bandwidth, we fill in the 3D {\bf Q}-space matrix by
filling all points from $\overrightarrow q_{min}$ to the 
$\overrightarrow q_{max}$, for all pixels in the face of the detector.
This quickly defines a volume of coverage, as shown schematically in
 two dimensions in Figure \ref{fig:ewald_fig}.

To reduce memory usage yet allow for quickly trying different detector
configurations, the 3D matrix of each sample orientation is saved a 32- or
64-bit integer, with each bit representing a particular detector. A simple
binary mask can then be applied to quickly compute the coverage if some
detectors are disabled, for example.

\subsection{Crystal Parameters and Single Reflection Coverage}

Crystal lattice parameters (lengths $a$, $b$, $c$ and angles $\alpha$, $\beta$,
$\gamma$) are typed in or loaded from the ISAW UB matrix file. CrystalPlan then
generates a list of $hkl$ peaks with d-spacings larger than the $d_{min}$
specified. For each sample orientation being simulated and for each peak, a function calculates at
what wavelength $\lambda$ and in which direction it scatters. If $\lambda$ 
is within range of the limits of the instrument, the beam direction is projected
onto each detector face to determine where, on each detector surface, it will be
measured.

Each $hkl$ peak is a separate object that holds a list of predicted
measurements, including which detector measured it, at what position on the
detector face, at what wavelength, and for which sample orientation. These data
will be used to  graphically display coverage and calculate statistics such as
redundancy.

\subsection{Crystal Symmetry}

When entering the lattice parameters, the crystal's point group symmetry can
also be entered. This information is then used to calculate the increased
measurement redundancy \cite{giacovazzo92}. A list of $n$ $3\times3$
multiplication matrices is generated based on the point group; any $hkl$ vector can be multiplied by
these matrices to find the $n$ equivalent $hkl$ reflections; for example, for
orthorombic (mmm) symmetry, $n=8$; therefore 8 multiplication matrices are
generated that multiply $hkl$ to give $\pm h$, $\pm k$, $\pm l$.

For the single reflections, each $hkl$ is tracked separately, but a primary
$hkl$ is found that points to a list of equivalent $hkl$'s. The GUI can
then display either all $hkl$, ignoring crystal symmetry; or combine them to
display only part of the reflections. Primary $hkl$'s
are picked so that they make a contiguous block.

In volume coverage mode, a 4D array is generated where the first 3 dimensions
correspond to {\bf Q}-space, and the $4^{th}$ dimension is a list of $n$
entries. Each entry in the $4^{th}$ dimension points to a voxel that is an
equivalent due to crystal symmetry. This allows CrystalPlan to quickly compute a volume coverage
map that takes into account crystal symmetry: at each voxel, all $n$ voxels are
checked to see if any of them were measured.

\subsection{Automatic Coverage Optimizer}

CrystalPlan allows the user to manually enter lists of sample orientations to
simulate; however, finding an optimal plan by trial and error would be very time
consuming. To make the process easier, CrystalPlan includes an Automatic
Coverage Optimizer, which takes as its inputs the desired number of sample
orientations and uses a genetic algorithm to find a solution that satisfies a
user-specified coverage criterion (for example, the user may ask for $85\%$ of
peaks to be measured, or for $90\%$ of {\bf Q}-volume).

Genetic algorithms have been covered in the literature before, see e.g
\cite{Goldberg89}. In this application, the genes
consist of goniometer angles. For instance, if a goniometer has freedom in 3 angles and the user
requests a 10 different sample orientations, then each individual has 30 genes.
The initial population is created randomly and is made to evolve using mutations
and crossover. Individuals whose goniometer angles turn out to be impossible to
reach are eliminated. For each individual set of sample orientations, the
coverage statistics are computed (taking crystal symmetry into account if desired), and this is used as the fitness of the given individual in the normal genetic
algorithm process.

The Automatic Coverage Optimizer can find an acceptable solution (if any is
possible) in times ranging from a few seconds to several minutes, depending on
the difficulty of the problem and the computing hardware. Given that at a
neutron facility, a measurement of a single orientation may take a few
hours, this feature can result in significant beam-time savings.

%---------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------
\section{Implementation}

\subsection{Language and Libraries}

The program needs to be cross-platform, and run equally well on Linux, Mac OS
and Windows, although Linux is our primary target operating system. Another
design goal was to use of open-source software and toolkits as much as possible.

To this end, CrystalPlan was written in Python \cite{python}, a powerful, open,
cross-platform scripting language. Python was chosen for its ease of cross-platform deployment, and for
its attractive language features. The numpy and scipy libraries (two
open-source, commonly installed Python libraries) were used for many of the
calculations \cite{numpy,scipy}. Some of the critical calculations were
also written with inline C code, which speed up some calculations by up to $400\times$ as compared with
the pure Python version (which can still be used, in case of incompatibilities).

\subsection{Code Layout}

The calculation code of CrystalPlan is placed in a model layer and kept
completely separate from eventual GUI elements, using a Model-View-Controller
architecture \cite{Gamma95}. This makes it possible to run calculations using
Python scripts, if desired. Unit tests for each part of the calculations are used to ensure
consistent results on different platforms and  systems.

\subsection{Graphical User Interface}

\subsubsection{Design.}

An attractive and easy-to-use GUI was considered an important part of
CrystalPlan, since the program is meant to be used interactively as a guide to
the experimenter. The user interface was written using wxPython
\cite{wxPython}, a mature cross-platform GUI toolkit, as well as the Enthought
Traits GUI. Some of the 2D plots use matplotlib \cite{matplotlib},
and the 3D visualization elements use the Mayavi2 toolkit for Python
\cite{mayavi}, itself based on VTK \cite{VTK}. All of these libraries are
open-source.

\subsubsection{Main Program Workflow.}
A screenshot of the main window GUI is shown in Figure \ref{fig:screenshots}.
The main window consists of a series of tabs, with workflow proceeding
intuitively from left to right:

\begin{itemize}
  	\item {\bf Q}-Space tab: define the volume of interest (minimum d-spacing)
  	and resolution of reciprocal space. 
 	\item Detectors tab: load instrument detector geometry
files, and enable or disable individual detectors. 
	\item Goniometer tab: to choose the
goniometer and degrees of orientation freedom of the instrument.
  	\item Sample tab: enter or load the crystal's lattice parameters and UB
  matrix.
  	\item Try an Orientation tab: interactively rotate a sample and observe the
  changes in coverage in real-time.
  	\item Add Orientations tab: type in a list of sample orientation angles to
  add to the experiment plan.

  	\item Experiment Plan tab: a list of the previously calculated sample
orientations. Here you can manage the orientations you will use in the experiment: add,
delete, or temporarily disable them to see the effect on coverage. Once
complete, the plan can be sent to the SNS data acquisition computers in order to
begin a run \cite{SNSDAS}.

\end{itemize}

\subsubsection{Reciprocal Space Coverage Visualization.}
Figure \ref{fig:volume_view}a) shows a screenshot of the reciprocal space
coverage 3D interface. The 3D view can represent {\bf Q}-space in two ways: volume coverage, or single-peak
coverage. The view shown can be rotated, translated or zoomed using the mouse or
keyboard.

The volume coverage shows a solid isosurface representing the volume measured at
least once. Redundant regions can also be shown using semi-transparent
isosurfaces; or the coverage can be inverted, showing non-measured volumes only;
or sliced between user-selectable $q_{min}$ and $q_{max}$.

The reflections view (Figure \ref{fig:refl_view}a) displays each $hkl$ as a
small sphere; the color of which indicates the number of times the $hkl$ is
predicted to be measured in the current experiment plan.

Both views can take into crystal symmetry in their display using the techniques
described previously. 
The volume coverage (Figures \ref{fig:volume_view}b) fills in the sphere of
coverage; 
whereas the single reflection view (Figure \ref{fig:refl_view}b) leaves only
the primary reflections (one eight of the total, in this example).

Finally, both views display quick coverage statistics in the lower right
corner, showing the percentage of volume (or of the number of peaks) measured,
as well as the proportion of redundancy (voxels or peaks measured 2 or more
times).

\subsubsection{Single Reflection View.}
Figure \ref{fig:single_reflection} shows a screenshot of the GUI allowing one to
observe in detail the predicted measurements of a particular peak, by clicking
in the 3D view or by typing in the $hkl$ values. The position on the detector
and wavelength of measurement are displayed. An additional window (not shown) allows the user to place the peak on a particular
position on a detector by automatically finding the necessary sample orientation
angles. This is used, for example, to center important peaks in the most
sensitive part of a detector.

%---------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------

\section{Sample Use Case}

As part of the TOPAZ beamline commissioning process, a Natrolite
(\ce{Na2Al2Si3O10} $\cdot$ \ce{2H2O}) single crystal was measured on a fixed
$45\degree\ \chi$ goniometer at room temperature and ambient atmosphere. An
initial short run (wavelength range from $\sim0.5$ \ang\ to $3.5$ \ang)
was processed using ISAW to determine the UB matrix and cell parameters of the crystal.

The resulting UB matrix was loaded into CrystalPlan, where the Automatic
Coverage Optimizer was used to generate an experiment plan with 8 sample
orientations (Figure \ref{fig:natrolite_gui}a) that covered 97\% of the peaks,
taking into account the crystal symmetry (orthorombic or mmm) (Figure
\ref{fig:natrolite_gui}b). The plan was left to run unattended overnight.

Peaks were integrated using ISAW, and the resulting .integrate file was loaded
into CrystalPlan for comparison. The predicted and actual measurement
positions on each detector face were compared for the 554 integrated peaks
found. Very good agreement was found; the
RMS difference between the predicted and actual measurement wavelength was only
0.016 \ang. The offsets in the peak position were plotted, which
helped show that the main contributor to this error was a slight misalignment of
the goniometer.

%, whereas the RMS of the detector position error was 7.6 mm.

%NOTE: Check on this misalignment of goniometer business - is that really the
% cause?

%---------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------
%---------------------------------------------------------------------------------------

\section{Conclusions and Future Development}

We have described CrystalPlan, a software program designed to quickly produce
efficient experiment plans for wide bandwidth time-of-flight Laue
diffractometers. This program will help increase scientific productivity at SNS
beamlines such as TOPAZ.

One planned future development for CrystalPlan include extending the software
to help predict coverage in inelastic neutron scattering experiments. Another
desired feature is the introduction of feedback loops to help determine the
right time to end data acquisition at a particular sample orientation.
CrystalPlan would interface with data analysis software such as ISAW and use
recently developped capabilities for real-time data analysis to monitor
the  signal to noise ratio on particular peaks. When the signal becomes
sufficient, the sample will automatically be moved to the next sample
orientation, further minimizing data acquisition time. Combined with advanced
sample changers coming on-line at TOPAZ, this will allow for extremely
efficient, unattended data acquisition.

     % Appendices appear after the main body of the text. They are prefixed by
     % a single \appendix declaration, and are then structured just like the
     % body text.

%\appendix
%\section{Appendix title}

%Text text text text text text text text text text text text text text
%text text text text text text text.

     %-------------------------------------------------------------------------
     % The back matter of the paper - acknowledgements and references
     %-------------------------------------------------------------------------

     % Acknowledgements come after the appendices

\ack{Acknowledgements}

This research is supported by UT Battelle, LLC under Contract No.
DE-AC05-00OR22725 for the U.S. Department of Energy, Office of Science.

     % References are at the end of the document, between \begin{references}
     % and \end{references} tags. Each reference is in a \reference entry.
\bibliographystyle{iucr}     
\bibliography{crystalplan}

%\begin{references}
%\reference{Author, A. \& Author, B. (1984). \emph{Journal} \textbf{Vol}, 
%first page--last page.}
%\end{references}

 %-------------------------------------------------------------------------
 % TABLES AND FIGURES SHOULD BE INSERTED AFTER THE MAIN BODY OF THE TEXT
 %-------------------------------------------------------------------------

 % Simple tables should use the tabular environment according to this
 % model

 % Postscript figures can be included with multiple figure blocks

\begin{figure}
\caption{2D Ewald sphere construction showing reciprocal-space detector coverage
for a wide-bandwidth Laue diffractometer. The smaller blue circle represents the
$1/\lambda_{max}$ limit of measurement, with the larger green circle
representing $1/\lambda_{min}$. The origin of reciprocal space (000) is common to both
spheres, but the ``virtual" sample position in the construction (S1 or S2)
depends on $\lambda$. The angular coverage of the detector is represented by the dashed
wedge, and projects an arc onto each sphere of radius $1/\lambda$. The
red wedge between the two $\lambda$ is built by projecting straight lines between the
$q_{min}$ and $q_{max}$ limits for each pixel on the detector face, giving the
volume in reciprocal space that is measured at a single sample oriention. }

\ifdograph
  \includegraphics[width=3.5in]{ewald_fig.ps}
\fi

\label{fig:ewald_fig} 
\end{figure}

\begin{figure}
\caption{Screenshot of the main GUI window. Six more interface windows (not
shown) are presented to the user when he or she clicks on the tabs at the
top of the main window.}
\ifdograph
  \includegraphics{main_gui}
\fi
\label{fig:screenshots}
\end{figure}

\begin{figure}
\caption{Sample screenshot showing the 3D view of the predicted coverage of
reciprocal space volume for 3 sample orientations and 14 detectors in TOPAZ, from 0.9 \ang\ to 4.0\ang:
a) without crystal symmetry, 17\% coverage; b) with crystal symmetry (mmm 8-fold
symmetry), coverage increases to 65\%. }
\ifdograph
  \includegraphics{volume_nosym}
\fi
a)
\ifdograph
  \includegraphics{volume_sym}
\fi
b)
\label{fig:volume_view} 
\end{figure}

\begin{figure}
\caption{Sample screenshot showing the 3D view of single-crystal reflection
measurements. Each sphere represents an $hkl$ peak, color-coded to indicate the
number of measurements (dark blue: 0, cyan: 1, green: 2, etc.). 
a) without crystal symmetry, 2375 of 9481 reflections are measured, or 25\%.
b) with crystal symmetry, 1276 of 1427 primary reflections are measured, or
89\%, and redundancy is also increased. }
\ifdograph
  \includegraphics{refl_nosym} 
\fi
a)
\ifdograph
  \includegraphics{refl_sym}
\fi
b)
\label{fig:refl_view}
\end{figure}

\begin{figure}
\caption{GUI showing details of the predicted measurement positions for a
single $hkl$.}
\ifdograph
  \includegraphics[width=2.5in]{single_reflection_3}
\fi

\label{fig:single_reflection} 
\end{figure}

\begin{figure} 
\caption{a) Screenshot of the Natrolite experiment plan generated using the
Automatic coverage Optimizer. b) Screenshot of the resulting coverage
statistics.}
\ifdograph
  \includegraphics[width=3.5in]{Plan8List}
\fi
a)
\ifdograph
  \includegraphics[width=3.5in]{NatCoverage}
\fi
b)
\label{fig:natrolite_gui}
\end{figure}

\end{document}                    % DO NOT DELETE THIS LINE
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%